---
title: "Preparing_dataset"
author: "Daniel Christiadi"
date: "04/10/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis', collapse = TRUE)
libraries <- c('readxl', 'tidyverse','fasttime', 'broom', 'outliers', 'skimr', 
               'plotly', 'heatmaply', 'shiny', 'data.table', 'reticulate',
               'gt', 'gtsummary')
installed <- installed.packages()[,'Package']
new.libs <- libraries[!(libraries %in% installed)]
if(length(new.libs)) install.packages(new.libs,repos="http://cran.csiro.au",
                                      dependencies=TRUE)
lib.ok <- sapply(libraries, require, character.only=TRUE)
thresh_chisq <<- 0.999999999

extraction_date <- fastPOSIXct("2021-05-01")
```

```{r Parsing raw data, include=FALSE, eval=FALSE}
lab_results <- fread(file = 'parsed_labresults.tsv', 
                     sep = '\t')

setnames(lab_results, 'PatientObjectID', 'id')
setnames(lab_results, 'Date_Local', 'date')
setnames(lab_results, 'ResultName', 'test1')
setnames(lab_results, 'OtherResultName', 'test2')
setnames(lab_results, 'Result', 'value')
setnames(lab_results, 'Units', 'unit1')
setnames(lab_results, 'OtherUnits', 'unit2')

pathology <- lab_results %>% 
    select(id, date, test1, test2, value, unit1, unit2) %>% 
    mutate(test = case_when(test1 == 'Other' ~ test2,
                            TRUE ~ test1)) %>% 
    mutate(unit = case_when(is.na(unit1) ~ unit2,
                            TRUE ~ unit1)) %>% 
    select(id, date, test, value, unit) %>% 
    mutate(value = parse_number(value))

# fwrite(pathology, file = 'pathology.gz')
```

```{r Extracting pathology data, include=FALSE}
pathology <- fread('pathology.gz')

# test name Glucose in mmol/L, serum and HgbA1c (Serum) in %
glucose <- pathology %>% 
    filter(test %in% c('Glucose, serum', 'Glucose')) %>% 
    select(id:value)

glucose$test <- 'glucose'

hba1c <- pathology %>% 
    filter(test == 'HgbA1c (Serum)') %>% 
    filter(value < 30) %>% # removed unrealistic result with no unit n = 2 
    select(id:value)

hba1c$test <- 'hba1c'

# creating diagnosis of diabetes based on random serum glucose >= 11.1 mmol/L and HbA1C > 6.5%
dt <- glucose %>% 
    mutate(diagnosis = case_when(value >= 11.1 ~ 'dkd',
                                 TRUE ~ 'no')) %>% 
    filter(diagnosis == 'dkd') %>% 
    distinct(across(c(id, diagnosis)))

dt2 <- hba1c %>% 
    mutate(diagnosis = case_when(value >= 6.5 ~ 'dkd',
                                 TRUE ~ 'no')) %>% 
    filter(diagnosis == 'dkd') %>% 
    distinct(across(c(id, diagnosis)))

dt3 <- pathology %>% 
    filter(test == '2Hr Glucose') %>% 
    mutate(diagnosis = case_when(value >= 11.1 ~ 'dkd',
                                 TRUE ~ 'no')) %>% 
    filter(diagnosis == 'dkd') %>% 
    distinct(across(c(id, diagnosis)))

dkd <- rbind(dt, dt2, dt3) %>% 
    distinct(across(c(id, diagnosis))) 

setnames(dkd, old = 'diagnosis', new = 'dkd')
rm(dt, dt2, dt3)

# test name Haemoglobin (g/L)
hb <- pathology %>% 
    filter(test %in% c('Haemoglobin','HAEMOGLOBIN')) %>% 
    filter(value <= 300) %>% # removing unrealistic Hb = 1187 g/L
    select(id:value)

hb[id == 19924 & value == 12.3]$value <- 123
hb$test <- 'haemoglobin'

# test name Haematocrit (%) : Haematocrit can be calculated (RBC*MCV)/10
hct <- pathology %>% 
    filter(test %in% c('Hct', 'Haematocrit')) %>% 
    select(id:value)

hct$test <- 'haematocrit'

# test name Red Blood Cell Count in x10^12/L
rbc <- pathology %>% 
    filter(test == 'Red Blood Cell Count') %>% 
    select(id:value)

rbc$test <- 'rbc'

# test name Red Cell Distribution Width. (%)
rdw <- pathology %>% 
    filter(test == 'Red Cell Distribution Width.') %>% 
    select(id:value)

rdw$test <- 'rdw'

# test name MCH in pg - MCH is calculated Hb (g/L) divided by RBC(in millions/microL)
mch <- pathology %>% 
    filter(test %in% c('Mean Corp. Haem.', 'MCH')) %>% 
    select(id:value)

mch$test <- 'mch'

# test name MCHC in g/L -  MCHC is calculated by Hb / Hct
mchc <- pathology %>% 
    filter(test == 'MCHC') %>% 
    select(id:value)

mchc$test <- 'mchc'

# test name MCV if fL
mcv <- pathology %>% 
    filter(test %in% c('Mean Cell Volume', 'MCV')) %>% 
    select(id:value)

mcv$test <- 'mcv'

# test name White Blood Cell Count in 10^9/L
wcc <- pathology %>% 
    filter(test %in% c('White Cell Count', 'White Blood Cell Count')) %>% 
    select(id:value)

wcc$test <- 'wcc'

# test name Actual Neutrophils (10^9/L)
neut <- pathology %>% 
    filter(test %in% c('Total Neuts', 'Actual Neutrophils')) %>% 
    select(id:value)

neut$test <- 'neutrophils'

# test name Actual Lymphocytes (10^9/L)
lymph <- pathology %>% 
    filter(test %in% c('Lymphocytes', 'Actual Lymphocytes'))

# changing unit x10^6/L to x10^9/L n = 2
lymph <- lymph %>% 
    mutate(value = case_when(unit %in% c('x10*6/L', 'x10^6/L') ~ value/1000,
                             TRUE ~ value))
lymph[unit == "%"]$value <- NA

lymph <- lymph %>% 
    select(id:value) 

lymph$test <- 'lymphocytes'

# test name Actual Eosinophils (10^9/L)
eosin <- pathology %>% 
    filter(test == 'Actual Eosinophils') %>% 
    select(id:value)

eosin$test <- 'eosinophils'

# test name Actual Monocytes (10^9/L)
monocyte <- pathology %>% 
    filter(test == 'Actual Monocytes') %>% 
    select(id:value)

monocyte$test <- 'monocytes'

# test name Actual Basophils (10^9/L)
baso <- pathology %>% 
    filter(test == 'Actual Basophils') %>% 
    select(id:value)

baso$test <- 'basophils'

# test name Platelet Count (10^9/L)
platelet <- pathology %>% 
    filter(test %in% c('Platelets', 'Platelet Count')) %>% 
    select(id:value)

platelet$test <- 'platelet'

# test name Ferritin (Serum) (microg/L)
ferritin <- pathology %>% 
    filter(test == 'Ferritin (Serum)') %>% 
    select(id:value) %>% 
    filter(value <= 40000) # removal of likely data error n = 4

ferritin$test <- 'ferritin'

# test name Iron (Serum) in umol/L
iron <- pathology %>% 
    filter(test == 'Iron (Serum)') %>% 
    select(id:value)

iron$test <- 'iron'

# test name Transferrin (Serum) in umol/ - conversion rate from g/L to umol/L is 12.5628
transferrin <- pathology %>% 
    filter(test == 'Transferrin (Serum)') %>% 
    mutate(value = case_when(unit %in% c('g/l', 'g/L') ~ value*12.5628,
                             TRUE ~ value)) %>% 
    select(id:value)

transferrin$test <- 'transferrin'

# test name Transferrin Saturation (Serum) (%)
tsat <- pathology %>% 
    filter(test == 'Transferrin Saturation (Serum)') %>% 
    select(id:value)

tsat$test <- 'tsat'

# test name Ca, adj; Calcium, serum; (mmol/L)
ca <- pathology %>% 
    filter(test %in% c('Ca, adj')) %>% 
    select(id:value)

ca$test <- 'calcium'

# test name Phosphorus, serum (mmol/L)
phosph <- pathology %>% 
    filter(test == 'Phosphorus, serum')

phosph[unit == '1.58']$value <- 1.58 # salvaging data n = 1

phosph <- phosph %>% 
    select(id:value) 

phosph$test <- 'phosphate'

# test name Cax Phos, adj | Calcium Phosphorus product
# dt <- pathology %>% filter(test == 'Cax Phos, adj' | test == 'Calcium Phosphorus product')
dt <- bind_rows(ca, phosph) %>% 
    pivot_wider(id_cols = c(id,date), 
                names_from = test,
                values_from = value, 
                values_fn = mean)

caphosproduct <- dt %>% 
    mutate(value = calcium*phosphate) %>%
    mutate(test = "caphosproduct") %>% 
    select(id, date, test, value)

# test name PTH Intact in pmol/L
pth <- pathology %>% 
    filter(test %in% c('Intact PTH', 'PTH Intact')) %>% 
    mutate(value = case_when(unit == 'pg/ml' ~ value*0.106,
                             TRUE ~ value)) %>% 
    select(id:value)

pth$test <- 'pth'

# test name Alkaline Phosphatase (Serum) in U/L
alkphos <- pathology %>% 
    filter(test %in% c('Alkaline Phosphatase (Serum)', 'Alk. Phos.', 'ALKP')) %>%
    select(id:value)

alkphos$test <- 'alkphos'

# test name Magnesium (Serum) in mmol/L
mg <- pathology %>% 
    filter(test %in% c('Magnesium', 'Magnesium (Serum)')) %>% 
    filter(value <= 10) # removed n = 6 values with no unit with unrealistic mg level > 80

mg[unit == 'mmol/day']$value <- NA # remove incorrect unit

mg <- mg %>% 
    select(id:value) 

mg$test <- 'magnesium'

# test name Sodium, serum in mmol/L
sodium <- pathology %>% 
    filter(test %in% c('Sodium', 'Sodium, serum')) %>% 
    filter(value != 5.8, value != 1238) %>%  # removal n = 2 values with unrealistic Na level 5.8 and 1238
    select(id:value) 

sodium$test <- 'sodium'

# test name Potassium, serum in mmol/L
potassium <- pathology %>% 
    filter(test %in% c('Potassium', 'Potassium, serum')) %>% 
    filter(value <= 10) %>%  # removal n = 5 values with unrealistic K level > 10
    select(id:value) 

potassium$test <- 'potassium'

# test name Chloride, serum in mmol/L
chloride <- pathology %>% 
    filter(test %in% c('Chloride, serum', 'Chloride')) %>% 
    filter(value > 70) # removal n = 9 values with unrealistic level < 20

chloride[unit == 'mmol/day']$value <- NA

chloride <- chloride %>% 
    select(id:value)

chloride$test <- 'chloride'

# test name Bicarbonate (Serum) in mmol/L
bicarb <- pathology %>% 
    filter(test %in% c('HCO3 (mEQ/L)', 'Bicarbonate (Serum)')) %>% 
    select(id:value)

bicarb$test <- 'bicarbonate'

# test name Alanine Aminotransferase in U/L
alt <- pathology %>% 
    filter(test == 'Alanine Aminotransferase') %>% 
    select(id:value)

alt$test <- 'alt'

# test name Albumin (Serum) in g/L
alb <- pathology %>% 
    filter(test == 'Albumin (Serum)') %>% 
    select(id:value) %>% 
    filter(value %>% between(10, 80)) # removed further n = 3 unrealistic value alb < 5 g/L and > 100

alb$test <- 'albumin'

# test name Gamma Glutamyl Transferase (Serum) in U/L
ggt <- pathology %>% 
    filter(test == 'Gamma Glutamyl Transferase (Serum)') %>% 
    select(id:value)

ggt$test <- 'ggt'

# test name Bilirubin Total (Serum) in umol/L
bili <- pathology %>% 
    filter(test == 'Bilirubin Total (Serum)') %>% 
    select(id:value) 

bili$test <- 'bilirubin'

# test name Globulin (Serum) in g/L
globulin <- pathology %>% 
    filter(test == 'Globulin (Serum)') %>% 
    select(id:value)

globulin$test <- 'globulin'

# test name Protein (Serum) in g/L
protein <- pathology %>% 
    filter(test %in% c('Protein (Serum)','Total Protein')) %>% 
    select(id:value)

protein$test <- 'protein'

# test name Cholesterol (Serum) in mmol/L
cholesterol <- pathology %>% 
    filter(test == 'Cholesterol (Serum)') %>% 
    select(id:value)

cholesterol$test <- 'cholesterol'

# test name Triglycerides (Serum) in mmol/L
tg <- pathology %>% 
    filter(test == 'Triglycerides (Serum)') %>% 
    select(id:value)

tg$test <- 'triglycerides'

# test name Low Density Lipoprotein (Serum) in mmol/L
ldl <- pathology %>% 
    filter(test == 'Low Density Lipoprotein (Serum)') %>% 
    select(id:value)

ldl$test <- 'ldl'

# test name High Density Lipoprotein (Serum) in mmol/L
hdl <- pathology %>% 
    filter(test == 'High Density Lipoprotein (Serum)') %>% 
    select(id:value)

hdl$test <- 'hdl'

# test name 'Thyroid Stimulating Hormone in mIU/L
tsh <- pathology %>% 
    filter(test == 'Thyroid Stimulating Hormone') %>% 
    select(id:value)

tsh$test <- 'tsh'

# test name Creatinine (Serum) in umol/L
creatinine <- pathology %>% 
    filter(test == 'Creatinine (Serum)') %>% 
    mutate(value = case_when(
        (unit == 'mmol/L' & value <= 10) ~ value*1000,
        TRUE ~ value
    )) %>% 
    select(id:value)

creatinine$test <- 'creatinine'

# test name eGFR mL/min/1.73m2
egfr <- pathology %>% 
    filter(test == 'eGFR') %>% 
    filter(value <= 300) %>% # excluding n = 60 with unrealistic egfr value > 300
    select(id:value)

egfr$test <- 'egfr'

# test name Urate in umol/L
urate <- pathology %>% 
    filter(test %in% c('Urate', 'Uric Acid (Serum)')) %>% 
    mutate(value = case_when(
        unit %in% c('mmol/l', 'mmol/L') ~ value*1000,
        value < 10 ~ value*1000,
        TRUE ~ value
    )) %>% 
    select(id:value)

urate$test <- 'urate'

# test name Urea (serum) in mmol/L
urea <- pathology %>% 
    filter(test == 'Urea (serum)') %>% 
    select(id:value)

urea$test <- 'urea'

# test name C-Reactive Protein in mg/L
crp <- pathology %>% 
    filter(test %in% c('CRP', 'C-Reactive Protein')) %>% 
    select(id:value)

crp$test <- 'crp'

# test name Erythrocyte Sedimentation Rate in mm/hr
esr <- pathology %>% 
    filter(test %in% c('ESR', 'Erythrocyte Sedimentation Rate')) %>% 
    select(id:value)

esr$test <- 'esr'

# test name Protein/Creatinine Ratio and Protein Creatinine Ratio (Urine)
protcreat <- pathology %>% 
    filter(test %in% c('Protein/creatinine', 'Protein/Creatinine', 
                       'Protein/Creatinine Ratio',
                       'Protein Creatinine Ratio (Urine)')) %>% 
    mutate(value = case_when(
        unit == 'g/g' ~ value*100,
        TRUE ~ value
    )) %>% 
    select(id:value)

protcreat$test <- 'protcreat_ratio'

# test name Albumin/creatinine, Albumin/Creatinine, Albumin Creatinine Ratio (Urine), Albumin/Creatinine Ratio, Albumin/Creatinine Ratio (Urine)
albcreat <- pathology %>% 
    filter(test %in% c('Albumin/creatinine', 'Albumin/Creatinine', 
                       'Albumin/Creatinine Ratio', 
                       'Albumin/Creatinine Ratio (Urine)')) %>% 
    select(id:value)

albcreat$test <- 'albcreat_ratio'

# 24h proteinuria in g/24h
dt <- read_excel('Diagnosis.xlsx', 
                 sheet = 'Urine24hrProtein',
                 col_names = c('id', 'date', 'test.name', 'other.test.name', 
                               'result.name', 'other.result', 'result', 
                               'result.str', 'normal.range', 'units', 
                               'provider'),
                 col_types = c('numeric', 'date', 'skip', 'skip', 'text', 
                               'skip', 'numeric', 'text', 'skip', 'text', 
                               'text'),
                 skip = 1, 
                 na = 'NULL')

proteinuria <- dt %>% 
    mutate(value = parse_number(result.str)) %>% 
    mutate(value = case_when(
        units == 'mg' ~ result/1000,
        TRUE ~ value
    )) %>% 
    mutate(test = "timedprotein") %>% 
    select(id, date, test, value)
```

```{r Demographic data}
patient <- read_excel('Diagnosis.xlsx', sheet = 'PtID',
                      col_names = c('id', 'birth.year', 'gender'),
                      col_types = c('numeric', 'numeric', 'text'),
                      skip = 1, na = 'NULL')
patient <- as.data.table(patient)
setkey(patient, id)
dt <- read_excel('need_gender_clarification.xlsx', sheet = 'Sheet1',
                 col_names = c('id', 'sex'),
                 col_types = c('numeric', 'text'),
                 skip = 1, na = 'NULL')
dt <- as.data.table(dt)
setkey(dt, id)
patient <- dt[patient]
patient[, gender := fifelse(gender %chin% c('Indeterminate', 'Unknown'), sex, gender)]
patient[, sex := NULL]
setkey(patient, id)
```

```{r Glomerulonephritis}
gn <- read_excel('Diagnosis.xlsx', sheet = 'GN Diagnosis',
                 col_names = c('id', 'diagnosis', 'date'),
                 col_types = c('numeric', 'text', 'date'), skip = 1,
                 na = 'NULL')
gn <- as.data.table(gn)
excluded_gn <- c('Acute systemic lupus erythematosus', 'Autoimmune vasculitis', 
                 'Cerebral vasculitis', 'Cryoglobulinemic vasculitis', 
                 'Cutaneous lupus erythematosus', 
                 'Cytomegalovirus-induced glomerulonephritis', 
                 'Diabetic glomerulonephritis', 'Discoid lupus erythematosus',
                 'Glomerulosclerosis', 
                 'Hemolytic anemia associated with systemic lupus erythematosus',
                 'Hyperfiltration focal segmental glomerulosclerosis', 
                 'Localized cutaneous vasculitis',
                 'Lupus erythematosus', 'Polyarteritis nodosa', 
                 'Primary systemic vasculitis', 
                 'Retinal vasculitis', 'Rheumatoid vasculitis', 
                 'Rupture of anterior cruciate ligament',
                 'Systemic lupus erythematosus', 
                 'Systemic lupus erythematosus arthritis', 
                 'Systemic lupus erythematosus-associated antiphospholipid syndrome', 
                 'Systemic lupus erythematosus with organ or system involvement', 
                 'Systemic lupus erythematosus, unspecified', 
                 'Urticarial vasculitis', 
                 'Vasculitis', 'Vasculitis secondary to drug ')
gn[, diagnosis := fifelse(diagnosis %chin% excluded_gn, 'no', diagnosis)]
gn <- gn[diagnosis != 'no',]
gn[, date := NULL]
lupus <- c('Systemic lupus erythematosus glomerulonephritis syndrome', 
          'Systemic lupus erythematosus glomerulonephritis syndrome, World Health Organization (WHO) class III',
          'Systemic lupus erythematosus glomerulonephritis syndrome, World Health Organization (WHO) class IV',
           'Systemic lupus erythematosus glomerulonephritis syndrome, World Health Organization (WHO) class V',
           'Systemic lupus erythematosus glomerulonephritis syndrome, World Health Organization (WHO) class VI')
unspecific <- c('Acute glomerulonephritis', 'Chronic diffuse glomerulonephritis', 
                'Chronic focal glomerulonephritis', 'Chronic glomerulonephritis',
                'Crescentic glomerulonephritis', 
                'Focal AND segmental proliferative glomerulonephritis',
                'Glomerulonephritis', 'Immune-complex glomerulonephritis', 
                'Necrotizing glomerulonephritis',
                'Proliferative glomerulonephritis', 'Renal vasculitis', 
                'Vasculitis secondary to drug')
mp_mcgn <- c('Acute nephritic syndrome, diffuse mesangial proliferative glomerulonephritis', 'C3 Glomerulonephritis', 
             'Mesangial proliferative glomerulonephritis', 
             'Mesangial proliferative glomerulonephritis', 
             'Mesangiocapillary glomerulonephritis', 
             'Mesangiocapillary glomerulonephritis, type I', 
             'Mesangiocapillary glomerulonephritis, type II')
pign <- c('Acute post-streptococcal glomerulonephritis', 
          'Post-infectious glomerulonephritis')
anca <- c('Eosinophilc Granulomatosis with Polyangiitis', 
          'Granulomatous Polyangiitis', 'Microscopic polyangiitis', 
          'Microscopic Polyangiitis', 
          'Primary pauci-immune necrotizing and crescentic glomerulonephritis',
          "Wegener's granulomatosis")
fg_itg <- c('Fibrillary glomerulonephritis', 'Immunotactoid Glomerulopathy')
fsgs <- c('Focal Segmental Glomerulosclerosis')
iga <- c('IgA nephropathy', 'Primary IgA nephropathy')
membranous <- c('Membranous glomerulonephritis', 
                'Membranous glomerulonephritis - stage III',
                'Nephrotic syndrome, diffuse membranous glomerulonephritis')
mcd <- c('Minimal change disease', 
         'Steroid-dependent minimal change glomerulonephritis')
gn_diagnosis <- list(lupus, unspecific, mp_mcgn, pign, anca, fg_itg, fsgs, iga, membranous, mcd)
names(gn_diagnosis) <- c('lupus nephritis', 'unspecific', 'mp/mcgn', 'pign', 'anca','fgn/itg', 'fsgs', 'igan', 'membranous',
                         'mcd')

for (i in seq_along(gn_diagnosis)){
  for(num in seq_along(gn$diagnosis)){
    if (gn$diagnosis[num] %in% gn_diagnosis[[i]]){{
      gn$diagnosis[num] <- names(gn_diagnosis[i])
    }}
  }
}

setkey(gn, id)
gn <- unique(gn)
# fixing patients with multiple diagnosis - diagnosis chosen based on the primary diagnosis
gn[id == 4339 & diagnosis == 'mcd', diagnosis := NA]
gn[id == 4551 & diagnosis == 'membranous', diagnosis := NA]
gn[id == 10143 & diagnosis == 'igan', diagnosis := NA]
gn[id == 15613 & diagnosis == 'mcd', diagnosis := NA]
gn[id == 18367 & diagnosis == 'unspecific', diagnosis := NA]
gn <- na.omit(gn)

```

```{r Vital signs}
dt <- read_excel('Diagnosis.xlsx', sheet = 'Exams', 
                 col_names = c('id', 'date', 'height', 'weight', 'bmi', 
                               'sit_dia', 'sit_sys', 'sit_hr', 'sta_dia', 
                               'sta_sys', 'sta_hr'),
                 col_types = c('numeric', 'date', 'numeric', 'numeric', 
                               'numeric', 'numeric', 'numeric', 'numeric', 
                               'numeric', 'numeric', 'numeric'),
                 skip = 1, na = 'NULL')
dt <- as.data.table(dt)
dt[, date := fastPOSIXct(date)]
dt[, diff_dia := sit_dia-sta_dia] 
dt[, diff_sys := sit_sys-sta_sys]
dt[, diff_hr := sit_hr-sta_hr]

exam <- melt(dt, id.vars = c('id', 'date'), na.rm = TRUE)
exam[, outlier := scores(value, type = 'chisq', prob = thresh_chisq)] 
exam <- exam[outlier == FALSE,]
exam <- exam[, outlier := NULL]
setnames(exam, old = 'variable', new = 'test')

# creating hypertension diagnosis 
dt2 <- exam[test == 'sit_sys' | test == 'sta_sys']
dt2[, htn := fifelse(value >= 140, 'htn', 'no')]
dt2 <- dt2[htn == 'htn',] 
setkey(dt2, id)

dt3 <- exam[test == 'sit_dia' | test == 'sta_dia']
dt3[, htn := fifelse(value >= 90, 'htn', 'no')]
dt3 <- dt3[htn == 'htn']
setkey(dt3, id)

htn <- rbind(dt2, dt3) %>% unique(by = 'id')
htn[, c('date', 'test', 'value') := NULL]
rm(dt2, dt3)
```

```{r Mortality}
death <- read_excel('Diagnosis.xlsx',
                    sheet = 'Death', 
                    col_names = c('id', 'date'),
                    col_types = c('numeric', 'date'), 
                    na = 'NULL', skip = 1)
death <- as.data.table(death)
setkey(death,id,date)
```

```{r Modality}
modality <- read_excel('Diagnosis.xlsx', 
                       sheet = 'modality',
                       col_names = c('id','date','end.date','modality','setting'),
                       col_types = c('numeric', 'date', 'date', 'text', 'text'),
                       skip = 1, na = 'NULL')
modality <- as.data.table(modality)
setkey(modality, id, date)
transplant <- modality[modality == 'Transplant', 
                       .(date = min(date), modality), 
                       by = id]
dialysis <- modality[modality %in% c('HD','PD','Nocturnal HD','Haemodiafiltration','Haemofiltration'),
                     .(date = min(date), modality), 
                     by = id]

eskd_definition <- c('HD','PD','Nocturnal HD','Haemodiafiltration','Haemofiltration', 'Transplant')
eskd_before <- modality[modality %chin% eskd_definition,]
setkey(eskd_before, id, date)
eskd_before <- eskd_before[, .SD[1], by = id]
```

```{r Melted pathology}
melted <- na.omit(unique(rbindlist(
    list(glucose, hba1c, exam, hb, hct, rdw, mch, mchc, mcv, wcc, rbc,
         neut, lymph, eosin, monocyte, baso, platelet, ferritin, iron, 
         transferrin, tsat, ca, phosph, pth, caphosproduct, alkphos, mg, 
         sodium, potassium, chloride, bicarb, alt, alb, ggt, bili, globulin, 
         protein, cholesterol, tg, ldl, hdl, tsh, creatinine, egfr, urate, 
         urea, crp, esr, protcreat, proteinuria, albcreat))))
setkey(melted, id, date)
saveRDS(melted, file = 'melted.RDS')
```

```{r Pts with EUC}
dt <- melted[test =='egfr']
setkey(dt, id, date)
egfr.init <- dt[, .SD[1], by = id]
egfr.init$test <- 'egfr_init'
egfr.last <- dt[, .SD[.N], by = id]
egfr.last$test <- 'egfr_last'
egfr.hit <- dt[dt[, .I[which(value <= 15)]], .SD[1], by = id]
egfr.hit$test <- 'eskd'
egfr.melt <- na.omit(rbindlist(list(egfr.init, egfr.last, egfr.hit)))
dt <- dcast(egfr.melt, id ~ test, value.var = c('value', 'date'), sep = '_')
setkey(dt, id)

dt %>% 
    distinct(across(id)) %>% 
    count()
```
There was a total of 15,433 patients ID extracted from pathology with 5,085 excluded due to absence of kidney function.
```{r Pts with missing demographic}
dt <- merge(death, dt, all.y = TRUE, by = 'id')
setnames(dt, old ='date', new = 'date.death')
dt <- merge(patient, dt, all.y = TRUE, by = 'id')
dt %>% filter(is.na(birth.year)) %>% count()
```
There was 68 patients with incomplete demographic data.
```{r Pts with pre-existing ESKD}
dt <- dt[!is.na(birth.year),]
dt <- merge(gn, dt, all.y = TRUE, by = 'id')
setnames(dt, old = 'diagnosis', new = 'gn')
dt[, gn := fifelse(is.na(gn), 'no', gn)]
dt <- merge(dkd, dt, all.y = TRUE, by = 'id')
dt[, dkd := fifelse(is.na(dkd), 'no', dkd)]
dt <- merge(htn, dt, all.y = TRUE, by = 'id')
dt[, htn := fifelse(is.na(htn), 'no', htn)]
dt <- merge(eskd_before, dt, all.y = TRUE, by = 'id')
dt[, c('end.date', 'setting') := NULL]
setnames(dt, old = 'date', new = 'existing_date')

dt <- dt %>% 
    relocate(c('existing_date', 'modality'), .after = date_eskd) %>% 
    mutate(preexisting_eskd = existing_date <= date_egfr_init) %>% 
    replace_na(list(preexisting_eskd = FALSE))

dt %>% count(preexisting_eskd)
```
A total of 493 patients had underlying ESKD leaving 9,787 for further analysis
```{r Outcomes dataset}
dt <- dt %>% 
    filter(preexisting_eskd == FALSE) %>% 
    select(-preexisting_eskd)

dt[date_eskd > existing_date, date_eskd := existing_date]
dt[!is.na(date_eskd) & is.na(modality) & value_egfr_last > 17, date_eskd := NA]
dt[, c('existing_date', 'modality') := NULL]

outcomes_dt <- dt %>% 
    relocate(date.death, .after = date_eskd) %>% 
    mutate(age_init = year(date_egfr_init) - birth.year) %>% 
    relocate(age_init, .after = birth.year)

fwrite(outcomes_dt, 'outcome_data.gz')
```
```{r Features dataset}
setkey(outcomes_dt, id)
melted$relyear <- (melted$date - outcomes_dt[.(melted$id), date_egfr_init])/31557600 
setattr(melted$relyear,'units','years')

features_dt <- melted %>% 
    filter(relyear >= 0)

setkey(features_dt, id)
fwrite(features_dt, 'features_data.gz')
```

```{r}
options(scipen = 5000)

short_dt <- outcomes_dt %>% 
    mutate(egfr_y = (date_egfr_last - date_egfr_init)/31557600) %>% 
    mutate(eskd_y = (date_eskd - date_egfr_init)/31557600) %>% 
    mutate(death_y = (date.death - date_egfr_init)/365.25) %>% 
    mutate(censor_y = (extraction_date - date_egfr_init)/365.25)

setattr(short_dt$eskd_y, 'units', 'years')
setattr(short_dt$egfr_y, 'units', 'years')
setattr(short_dt$death_y, "units", "years")
setattr(short_dt$censor_y, "units", "years")

m <- features_dt[test == 'egfr'][, .N, by = 'id']
setkey(m, id)
setnames(m, old = 'N', new = 'egfr_freq')
short_dt <- merge(m, short_dt, all.y = TRUE, by = 'id')

short_dt %>% 
    mutate(egfr_freq_less_than_two = egfr_freq < 2) %>% 
    count(egfr_freq_less_than_two)
```
There was a total of 1,682 patients with eGFR frequency less than 2. They were excluded leaving 8,105 patients
```{r}
short_dt <- short_dt %>% 
    mutate(egfr_freq_less_than_two = egfr_freq < 2) %>% 
    filter(egfr_freq_less_than_two == FALSE) %>% 
    select(-egfr_freq, -birth.year, -value_egfr_last, -value_eskd, -egfr_freq_less_than_two) %>% 
    mutate(time = pmin.int(death_y, eskd_y, censor_y, na.rm = TRUE))

short_dt %>% 
    filter(time <= 0.1) %>% 
    count()
```
There were 537 patients who developed ESKD (521), death (6), or censored (10) less or equal than 0.1 years (36 days), which were, then, excluded. (There were 7,568 patients left. The mean time 0.00450 years (~1 day) and median time 0. 

```{r Excluded dataset}
excluded_dt <- short_dt %>% 
    filter(time <= 0.1) 
    
excluded_dt <- excluded_dt %>% 
    mutate(event = case_when(
        time == as.numeric(death_y) ~ "death",
        time == as.numeric(eskd_y) ~ "eskd",
        TRUE ~ "censored"
    ))
```
As there was only 1 case of PIGN, the decision was to collapse PIGN into unspecific GN diagnosis.
```{r Creating short_dt}
short_dt[gn == "pign", gn := "unspecific"]

short_dt <- short_dt %>% 
    filter(time > 0.1) %>% 
    mutate(event = case_when(
        time == as.numeric(death_y) ~ "death",
        time == as.numeric(eskd_y) ~ "eskd",
        TRUE ~ "censored"
    )) %>% 
    mutate(event = factor(event, 
                          levels = c("censored", "eskd", "death"))) %>% 
    select(-(date_egfr_init:censor_y)) %>% 
    mutate(htn = factor(htn, levels = c("no", "htn"))) %>% 
    mutate(dkd = factor(dkd, levels = c("no", "dkd"))) %>% 
    mutate(gn = factor(gn, 
                       levels = c("no", "anca", "fgn/itg", "fsgs", "igan", 
                                  "lupus nephritis", "mcd", "membranous",
                                  "mp/mcgn", "unspecific"))) %>% 
    mutate(gender = factor(gender))

saveRDS(short_dt, "short_dt.RDS")
```

```{r Creating long_dt}
long_dt <- features_dt %>% 
    filter(id %in% short_dt$id)

long_dt <- long_dt %>% 
    select(-date) %>% 
    mutate(relyear = round(relyear, digits = 1))

long_dt <- long_dt %>%
    pivot_wider(id_cols = c("id", "relyear"),
                names_from = test,
                values_from = value,
                values_fn = mean)

long_dt <- long_dt %>% 
    left_join(dplyr::select(short_dt, id, time), by = "id")

long_dt <- long_dt %>% 
    relocate(time, .after = relyear) %>% 
    filter(relyear < time) %>% 
    select(-time)

fwrite(long_dt, file = "long_dt.gz")
```

```{r Investigation of missing test}
miss_function <- function(dt_name, test_name){
   dt_name %>% 
    group_by(id) %>% 
    mutate(missing_test = sum(is.na({{test_name}}))) %>% 
    add_count(id) %>% 
    select(id, {{test_name}}, missing_test, n) %>% 
    filter(missing_test == n) %>% 
    select(id, missing_test, n) %>% 
    distinct(id, .keep_all = TRUE) 
}

egfr_miss <- miss_function(long_dt, egfr)
calcium_miss<- miss_function(long_dt, calcium)
phosph_miss <- miss_function(long_dt, phosphate)
bicarbonate_miss <- miss_function(long_dt, bicarbonate)
albumin_miss <- miss_function(long_dt, albumin)
magnesium_miss <- miss_function(long_dt, magnesium)
alkphos_miss <- miss_function(long_dt, alkphos)
sodium_miss <- miss_function(long_dt, sodium)
potassium_miss <- miss_function(long_dt, potassium)
chloride_miss <- miss_function(long_dt, chloride)
urate_miss <- miss_function(long_dt, urate)
pth_miss <- miss_function(long_dt, pth)
crp_miss <- miss_function(long_dt, crp)
hb_miss <- miss_function(long_dt, haemoglobin)
plt_miss <- miss_function(long_dt, platelet)
neut_miss <- miss_function(long_dt, neutrophils)
lymph_miss <- miss_function(long_dt, lymphocytes)
bmi_miss <- miss_function(long_dt, bmi)
glu_miss <- miss_function(long_dt, glucose)
diffsys_miss <- miss_function(long_dt, diff_sys)
diffdia_miss <- miss_function(long_dt, diff_dia)

esr_miss <- miss_function(long_dt, esr)

pcracr_miss <- long_dt %>% 
    group_by(id) %>% 
    mutate(missing_acr = sum(is.na(albcreat_ratio)),
           missing_pcr = sum(is.na(protcreat_ratio))) %>% 
    add_count(id) %>% 
    select(id, protcreat_ratio, albcreat_ratio, missing_acr, missing_pcr, n) %>%
    filter(missing_acr == n | missing_pcr == n) %>% 
    distinct(id, .keep_all = TRUE)

```

```{r}
patient <- readxl::read_excel("Diagnosis.xlsx", sheet = "PtID")
```





