---
title: "Preparing Dataset"
author: "Daniel Christiadi"
date: "Compiled on `r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis', collapse = TRUE)
libraries <- c('readxl', 'tidyverse', 'janitor', 'skimr', 'heatmaply', 
               'data.table', 'gtsummary', 'lubridate')
installed <- installed.packages()[,'Package']
new.libs <- libraries[!(libraries %in% installed)]
if(length(new.libs)) install.packages(new.libs,repos="http://cran.csiro.au",
                                      dependencies=TRUE)
lib.ok <- sapply(libraries, require, character.only=TRUE)
extraction_date <<- as_datetime("2022-01-14")
```

```{r parsing text data part 1, include=FALSE, eval=FALSE}
lab_results <- fread(file = 'parsed_labresults.tsv', 
                     sep = '\t')

setnames(lab_results, 'PatientObjectID', 'id')
setnames(lab_results, 'Date_Local', 'date')
setnames(lab_results, 'ResultName', 'test1')
setnames(lab_results, 'OtherResultName', 'test2')
setnames(lab_results, 'ResultString', 'value1')
setnames(lab_results, "Result", "value2")
setnames(lab_results, 'Units', 'unit1')
setnames(lab_results, 'OtherUnits', 'unit2')
setnames(lab_results, "MappedUnits", 'unit3')

pathology <- lab_results %>% 
    select(id, date, test1, test2, value1, value2, unit1, unit2, unit3) %>%
    mutate(test = case_when(
        test1 == "Other" ~ test2,
        TRUE ~ test1
    )) %>%
    mutate(unit = case_when(
        unit3 == "Other" & unit1 == "NULL" ~ unit2,
        unit3 == "Other" & unit1 == "*" ~ unit2,
        unit3 == "Other" ~ unit1,
        TRUE ~ unit3
    )) %>%
    mutate(value1 = parse_number(value1),
           value2 = parse_number(value2),
           value = if_else(is.na(value1), value2, value1)) %>% 
    select(id, date, test, value, unit)

# fwrite(pathology, "pathology.gz")
```

```{r parsing text data part 2, include=FALSE, eval=FALSE}
pathology <- fread('pathology.gz')

pathology <- pathology %>% 
    drop_na(value)

# test name Glucose in mmol/L, serum and HgbA1c (Serum) in %
glucose <- pathology %>% 
    filter(test == 'Glucose, serum') %>% 
    select(id:value)

glucose$test <- 'glucose'

hba1c <- pathology %>% 
    filter(test %in% c('HgbA1c (Serum)', "Hb A1C")) %>% 
    filter(value < 30) %>% # removed unrealistic result with no unit n = 2 
    select(id:value)

hba1c$test <- 'hba1c'

# test name Haemoglobin (g/L)
hb <- pathology %>% 
    filter(test %in% c('Haemoglobin','HAEMOGLOBIN')) %>% 
    filter(value <= 300) %>% # removing unrealistic Hb = 1187 g/L
    select(id:value)

hb[id == 19924 & value == 12.3]$value <- 123
hb$test <- 'haemoglobin'

# test name Haematocrit (%) : Haematocrit can be calculated (RBC*MCV)/10
hct <- pathology %>% 
    filter(test == 'Haematocrit') %>% 
    select(id:value)

hct$test <- 'haematocrit'

# test name Red Blood Cell Count in x10^12/L
rbc <- pathology %>% 
    filter(test == 'Red Blood Cell Count') %>% 
    select(id:value)

rbc$test <- 'rbc'

# test name Red Cell Distribution Width. (%)
rdw <- pathology %>% 
    filter(test %in% c('Red Cell Distribution Width.', "RDW")) %>%
    select(id:value)

rdw$test <- 'rdw'

# test name MCH in pg - MCH is calculated Hb (g/L) divided by RBC(in millions/microL)
mch <- pathology %>% 
    filter(test %in% c('Mean Corp. Haem.', 'MCH')) %>% 
    select(id:value)

mch$test <- 'mch'

# test name MCHC in g/L -  MCHC is calculated by Hb / Hct
mchc <- pathology %>% 
    filter(test == 'MCHC') %>% 
    select(id:value)

mchc$test <- 'mchc'

# test name MCV if fL
mcv <- pathology %>% 
    filter(test %in% c('Mean Cell Volume', 'MCV')) %>% 
    select(id:value)

mcv$test <- 'mcv'

# test name White Blood Cell Count in 10^9/L
wcc <- pathology %>% 
    filter(test == 'White Blood Cell Count') %>% 
    select(id:value)

wcc$test <- 'wcc'

# test name Actual Neutrophils (10^9/L)
neut <- pathology %>% 
    filter(test %in% c('Total Neuts', 'Actual Neutrophils')) %>% 
    select(id:value)

neut$test <- 'neutrophils'

# test name Actual Lymphocytes (10^9/L)
lymph <- pathology %>% 
    filter(test == 'Actual Lymphocytes') %>% 
    mutate(value = case_when(unit == 'x10*6/L' ~ value/1000,
                             TRUE ~ value)) %>% 
    select(id:value)

lymph$test <- 'lymphocytes'

# test name Actual Eosinophils (10^9/L)
eosin <- pathology %>% 
    filter(test == 'Actual Eosinophils') %>% 
    select(id:value)

eosin$test <- 'eosinophils'

# test name Actual Monocytes (10^9/L)
monocyte <- pathology %>% 
    filter(test == 'Actual Monocytes') %>% 
    select(id:value)

monocyte$test <- 'monocytes'

# test name Actual Basophils (10^9/L)
baso <- pathology %>% 
    filter(test == 'Actual Basophils') %>% 
    select(id:value)

baso$test <- 'basophils'

# test name Platelet Count (10^9/L)
platelet <- pathology %>% 
    filter(test %in% c('Platelets', 'Platelet Count')) %>% 
    select(id:value)

platelet$test <- 'platelet'

# test name Ferritin (Serum) (microg/L)
ferritin <- pathology %>% 
    filter(test == 'Ferritin (Serum)') %>% 
    select(id:value)

ferritin$test <- 'ferritin'

# test name Iron (Serum) in umol/L
iron <- pathology %>% 
    filter(test == 'Iron (Serum)') %>% 
    select(id:value)

iron$test <- 'iron'

# test name Transferrin (Serum) in umol/ - conversion rate from g/L to umol/L is 12.5628
transferrin <- pathology %>% 
    filter(test == 'Transferrin (Serum)') %>% 
    mutate(value = case_when(unit %in% c('g/l', 'g/L') ~ value*12.5628,
                             TRUE ~ value)) %>% 
    select(id:value)

transferrin$test <- 'transferrin'

# test name Transferrin Saturation (Serum) (%)
tsat <- pathology %>% 
    filter(test == 'Transferrin Saturation (Serum)') %>% 
    select(id:value)

tsat$test <- 'tsat'

# test name Ca, adj; Calcium, serum; (mmol/L)
ca_adj <- pathology %>% 
    filter(test == "Ca, adj") %>% 
    select(id:value)

ca <- pathology %>% 
    filter(test == "Calcium, serum") %>% 
    filter(value < 10, value != 0) %>% 
    select(id:value)

# test name Phosphorus, serum (mmol/L)
phosph <- pathology %>% 
    filter(test == 'Phosphorus, serum') %>% 
    filter(value < 10) %>% # removing likely false data 
    select(id:value)

phosph$test <- 'phosphate'

# test name PTH Intact in pmol/L
pth <- pathology %>% 
    filter(test %in% c('Intact PTH', 'PTH Intact')) %>% 
    mutate(value = case_when(unit == 'pg/ml' ~ value*0.106,
                             TRUE ~ value)) %>% 
    select(id:value)

pth$test <- 'pth'

# test name Alkaline Phosphatase (Serum) in U/L
alkphos <- pathology %>% 
    filter(test %in% c('Alkaline Phosphatase (Serum)', 'Alk. Phos.', 
                       'ALKP')) %>%
    select(id:value)

alkphos$test <- 'alkphos'

# test name Magnesium (Serum) in mmol/L
mg <- pathology %>% 
    filter(test == 'Magnesium (Serum)') %>%
    filter(value <= 10) %>% # removed n = 6 values with no unit 
    select(id:value)

mg$test <- 'magnesium'

# test name Sodium, serum in mmol/L
sodium <- pathology %>% 
    filter(test == 'Sodium, serum') %>%
    filter(value != 5.8, value != 1238) %>%  # removal n = 2 values with unrealistic Na level 5.8 and 1238
    select(id:value) 

sodium$test <- 'sodium'

# test name Potassium, serum in mmol/L
potassium <- pathology %>% 
    filter(test == 'Potassium, serum') %>%
    filter(value <= 10) %>% # removal n = 5 values with unrealistic K level > 10
    select(id:value) 

potassium$test <- 'potassium'

# test name Chloride, serum in mmol/L
chloride <- pathology %>% 
    filter(test == 'Chloride, serum') %>% 
    filter(value > 70) %>%  # removal n = 9 values with unrealistic level < 20
    select(id:value)

chloride$test <- 'chloride'

# test name Bicarbonate (Serum) in mmol/L
bicarb <- pathology %>% 
    filter(test == 'Bicarbonate (Serum)') %>% 
    select(id:value)

bicarb$test <- 'bicarbonate'

# test name Alanine Aminotransferase in U/L
alt <- pathology %>% 
    filter(test == 'Alanine Aminotransferase') %>% 
    select(id:value)

alt$test <- 'alt'

# test name Albumin (Serum) in g/L
alb <- pathology %>% 
    filter(test == 'Albumin (Serum)') %>% 
    select(id:value) %>% 
    filter(value %>% between(10, 80)) # removed further n = 3 unrealistic value alb < 5 g/L and > 100

alb$test <- 'albumin'

# test name Gamma Glutamyl Transferase (Serum) in U/L
ggt <- pathology %>% 
    filter(test == 'Gamma Glutamyl Transferase (Serum)') %>% 
    select(id:value)

ggt$test <- 'ggt'

# test name Bilirubin Total (Serum) in umol/L
bili <- pathology %>% 
    filter(test == 'Bilirubin Total (Serum)') %>% 
    select(id:value) 

bili$test <- 'bilirubin'

# test name Globulin (Serum) in g/L
globulin <- pathology %>% 
    filter(test == 'Globulin (Serum)') %>% 
    select(id:value)

globulin$test <- 'globulin'

# test name Protein (Serum) in g/L
dt <- pathology %>% 
    filter(test %in% c('Protein (Serum)','Total Protein')) %>% 
    pivot_wider(id_cols = c("id", "date"),
                names_from = test,
                values_from = value,
                values_fn = mean)
setnames(dt, old = "Total Protein", "total_protein")  
setnames(dt, old = "Protein (Serum)", "serum_protein")  

protein <- dt %>% 
    mutate(value = if_else(is.na(total_protein), 
                           serum_protein, total_protein),
           test = "protein") %>% 
    select(id, date, test, value)

# test name Creatinine (Serum) in umol/L
creatinine <- pathology %>% 
    filter(test == 'Creatinine (Serum)') %>% 
    mutate(value = case_when(
        (unit == 'mmol/L' & value <= 10) ~ value*1000,
        TRUE ~ value
    )) %>% 
    filter(value > 0) %>% 
    select(id:value)

creatinine$test <- 'creatinine'

# test name eGFR mL/min/1.73m2
egfr <- pathology %>% 
    filter(test == "eGFR") %>% 
    mutate(unit_factor = if_else(unit == "NULL", "from_null", 
                                 "from_nonnull")) %>% 
    pivot_wider(id_cols = c("id", "date"),
                names_from = unit_factor,
                values_from = value,
                values_fn = mean) %>% 
    mutate(value = if_else(is.na(from_nonnull), from_null, from_nonnull),
           test = "egfr") %>% 
    filter(value <= 90) %>% 
    select(id, date, test, value)

# test name Urate in umol/L
urate <- pathology %>% 
    filter(test %in% c('Urate', 'Uric Acid (Serum)')) %>% 
    mutate(value = if_else(unit %in% c('mmol/l', 'mmol/L'), value*1000, 
                           value),
           value = if_else(value < 10, value*1000, value)) %>% 
    select(id:value)

urate$test <- 'urate'

# test name Urea (serum) in mmol/L
urea <- pathology %>% 
    filter(test == 'Urea (serum)') %>% 
    mutate(unit_factor = if_else(unit == "NULL", "from_null", 
                                 "from_nonnull")) %>% 
    pivot_wider(id_cols = c("id", "date"),
                names_from = unit_factor,
                values_from = value,
                values_fn = mean) %>% 
    mutate(value = if_else(is.na(from_nonnull), from_null, from_nonnull),
           test = "urea") %>% 
    filter(value < 200) %>%
    select(id, date, test, value)

# test name C-Reactive Protein in mg/L
crp <- pathology %>% 
    filter(test %in% c('CRP', 'C-Reactive Protein')) %>%
    mutate(test = "crp") %>% 
    select(id, date, test, value)

# test name Erythrocyte Sedimentation Rate in mm/hr
esr <- pathology %>% 
    filter(test %in% c('ESR', 'Erythrocyte Sedimentation Rate', "E.S.R.")) %>% 
    mutate(test = "esr") %>% 
    select(id:value)

# test name Protein/Creatinine Ratio in mg/mmol
protcreat <- pathology %>% 
    filter(test %in% c('Protein/creatinine', 'Protein/Creatinine', 
                       'Protein/Creatinine Ratio',
                       'Protein Creatinine Ratio (Urine)')) %>% 
    mutate(value = if_else(unit == 'g/g', value*100, value),
           test = "protcreat_ratio") %>% 
    select(id, date, test, value)

protcreat$test <- 'protcreat_ratio'

# test name Albumin/Creatinine Ratio in mg/mmol
albcreat <- pathology %>% 
    filter(test %in% c('Albumin/creatinine', 'Albumin/Creatinine', 
                       'Albumin/Creatinine Ratio', 
                       'Albumin/Creatinine Ratio (Urine)')) %>% 
    mutate(test = "albcreat_ratio") %>% 
    select(id, date, test, value)

albcreat$test <- 'albcreat_ratio'
```

```{r parsing raw data, include=FALSE, eval=FALSE}
dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "diabetes",na = "NULL", skip = 1,
                 col_names = c("id", "date", "glucose", "hba1c"),
                 col_types = c("numeric", "date", "numeric", "numeric"))
dt <- dt %>% 
    arrange(id, date) %>% 
    mutate(hba1c = replace(hba1c, 
                           id == 22476 & hba1c == 87,
                           values = 8.7)) %>% 
    pivot_longer(cols = c("glucose", "hba1c"),
                 names_to = "test",
                 values_to = "value") %>% 
    drop_na(value)

glucose <- dt %>% 
    filter(test == "glucose") %>%
    full_join(y = glucose, by = c("id", "date")) %>% 
    rename(test_x = "test.x", value_x = "value.x",
           test_y = "test.y", value_y = "value.y") %>% 
    mutate(test_x = if_else(is.na(test_x), test_y, test_x),
           value_x = if_else(is.na(value_x), value_y, value_x)) %>% 
    select(id, date, test_x, value_x) %>% 
    rename(test = test_x, value = value_x) %>% 
    distinct() %>% 
    filter(value > 0)

hba1c <- dt %>% 
    filter(test == "hba1c") %>% 
    full_join(y = hba1c, by = c("id", "date")) %>% 
    rename(test_x = "test.x", value_x = "value.x",
           test_y = "test.y", value_y = "value.y") %>% 
    mutate(test_x = if_else(is.na(test_x), test_y, test_x),
           value_x = if_else(is.na(value_x), value_y, value_x)) %>% 
    select(id, date, test_x, value_x) %>% 
    rename(test = test_x, value = value_x) %>% distinct()

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "PCR", na = "NULL")
dt <- dt %>% 
    rename(id = PatientObjectID, date = ResultDate_Local,
           test = ResultName, value_num = ResultNumeric, value = Result) %>% 
    select(id, date, test, value_num, value) %>% 
    mutate(value = parse_number(value),
           test = "protcreat_ratio") %>% 
    select(-value_num)

protcreat <- dt %>% 
    full_join(y = protcreat, by = c("id", "date")) %>% 
    rename(test_x = "test.x", value_x = "value.x",
           test_y = "test.y", value_y = "value.y") %>% 
    mutate(test_x = if_else(is.na(test_x), test_y, test_x),
           value_x = if_else(is.na(value_x), value_y, value_x)) %>% 
    select(id, date, test_x, value_x) %>% 
    rename(test = test_x, value = value_x) %>% distinct()

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "ACR", na = "NULL")
dt <- dt %>% 
    rename(id = PatientObjectID, date = ResultDate_Local, test = ResultName,
           test2 = OtherResultName, value = Result, 
           value_num = ResultNumeric) %>% 
    mutate(test = if_else(test == "Other", test2, test),
           value = parse_number(value),
           test = "albcreat_ratio") %>% 
    select(id, date, test, value)

albcreat <- dt %>% 
    full_join(y = albcreat, by = c("id", "date")) %>% 
    rename(test_x = "test.x", value_x = "value.x",
           test_y = "test.y", value_y = "value.y") %>% 
    mutate(test_x = if_else(is.na(test_x), test_y, test_x),
           value_x = if_else(is.na(value_x), value_y, value_x)) %>% 
    select(id, date, test_x, value_x) %>% 
    rename(test = test_x, value = value_x) %>% distinct()

# converting protein/creatinine ratio to albumin/creatinine ratio. Current unit of uPCR is in mg/mmol. To convert the unit to mg/g, the value is multiplied by 8.85. In contrast to convert from mg/g to mg/mmol, the value is multiplied by 0.113
albcreat <- protcreat %>% 
    mutate(value_mgg = value*8.85) %>% 
    mutate(log_acr_mgg = case_when(
        value_mgg < 40 ~ 0.9518+0.1264*log(value_mgg),
        between(value_mgg, 40, 60, 
                incbounds = TRUE) ~ -1.2568+0.7251*log(value_mgg),
        between(value_mgg, 60, 250, 
                incbounds = TRUE) ~ -6.7837+2.0751*log(value_mgg),
        between(value_mgg, 250, 1000, 
                incbounds = TRUE) ~ -2.9649+1.3834*log(value_mgg),
        TRUE ~ -0.0239+0.9577*log(value_mgg)
    )) %>% 
    mutate(acr_mgg = exp(log_acr_mgg)) %>% 
    mutate(acr_value = acr_mgg*0.113) %>% 
    select(id, date, acr_value) %>% 
    full_join(albcreat, by = c("id", "date")) %>% 
    distinct() %>% 
    arrange(id, date) %>% 
    replace_na(list(test = "albcreat_ratio")) %>% 
    mutate(value = if_else(is.na(value),
                           acr_value,
                           value)) %>% 
    select(id, date, test, value)

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "haem1", na = "NULL") 
haem <- dt %>% 
    arrange(PatientObjectID, LabTestDate_Local) %>% 
    rename(id = PatientObjectID, date = LabTestDate_Local, 
           hct = HctNum, hb = HgbNum, rbc = RBCNum, rdw = RBWNum,
           mchc = MCHCNum, mch = MCHNum, mcv = MCVNum, wbc = WBCNum,
           plt = PlateletsNum)

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "haem2", na = "NULL") 
haem <- dt %>% 
    arrange(PatientObjectID, LabTestDate_Local) %>% 
    rename(id = PatientObjectID, date = LabTestDate_Local,
           neut = NeutrophilsNum, lymph = LymphsNum, eosin = EosinophilsNum,
           mono = MonosNum, baso = BasosNum) %>% 
    full_join(y = haem, by = c("id", "date"))

neut <- haem %>% 
    select(id, date, neut) %>% 
    full_join(y = neut, by = c("id", "date")) %>%
    mutate(neut = if_else(is.na(neut), value, neut)) %>% 
    select(id, date, neut) %>% 
    mutate(test = "neutrophils") %>% 
    rename(value = neut) %>% 
    relocate(test, .before = value) %>% 
    distinct() 

lymph <- haem %>% 
    select(id, date, lymph) %>% 
    full_join(y = lymph, by = c("id", "date")) %>% 
    mutate(lymph = if_else(lymph > 500, value, lymph),
           lymph = if_else(is.na(lymph), value, lymph)) %>% 
    select(id, date, lymph) %>% 
    mutate(test = "lymphocytes") %>% 
    rename(value = lymph) %>% 
    relocate(test, .before = value) %>% 
    distinct()

eosin <- haem %>% 
    select(id, date, eosin) %>% 
    full_join(y = eosin, by = c("id", "date")) %>% 
    mutate(eosin = if_else(is.na(eosin), value, eosin)) %>% 
    select(id, date, eosin) %>% 
    mutate(test = "eosinophils") %>% 
    rename(value = eosin) %>% 
    relocate(test, .before = value) %>% distinct()

monocyte <- haem %>% 
    select(id, date, mono) %>% 
    full_join(y = monocyte, by = c("id", "date")) %>% 
    mutate(mono = if_else(is.na(mono), value, mono)) %>% 
    select(id, date, mono) %>% 
    mutate(test = "monocytes") %>% 
    rename(value = mono) %>% 
    relocate(test, .before = value) %>% distinct()

baso <- haem %>% 
    select(id, date, baso) %>% 
    full_join(y = baso, by = c("id", "date")) %>% 
    mutate(baso = if_else(is.na(baso), value, baso)) %>% 
    select(id, date, baso) %>% 
    mutate(test = "basophils") %>% 
    rename(value = baso) %>% 
    relocate(test, .before = value) %>% distinct()

hct <- haem %>% 
    select(id, date, hct) %>% 
    full_join(y = hct, by = c("id", "date")) %>% 
    mutate(hct = if_else(is.na(hct), value, hct)) %>% 
    select(id, date, hct) %>% 
    mutate(test = "haematocrit") %>% 
    rename(value = hct) %>% 
    relocate(test, .before = value) %>% distinct()

hb <- haem %>% 
    select(id, date, hb) %>% 
    full_join(y = hb, by = c("id", "date")) %>%
    mutate(hb = if_else(is.na(hb), value, hb)) %>% 
    select(id, date, hb) %>% 
    mutate(test = "haemoglobin") %>%
    rename(value = hb) %>% 
    relocate(test, .before = value) %>% distinct()

rbc <- haem %>% 
    select(id, date, rbc) %>% 
    full_join(y = rbc, by = c("id", "date")) %>% 
    mutate(rbc = if_else(is.na(rbc), value, rbc)) %>% 
    select(id, date, rbc) %>% 
    mutate(test = "rbc") %>% 
    rename(value = rbc) %>% 
    relocate(test, .before = value) %>% distinct()

rdw <- haem %>% 
    select(id, date, rdw) %>% 
    full_join(y = rdw, by = c("id", "date")) %>% 
    mutate(rdw = if_else(is.na(rdw), value, rdw)) %>% 
    select(id, date, rdw) %>% 
    mutate(test = "rdw") %>% 
    rename(value = rdw) %>% 
    relocate(test, .before = value) %>% distinct()

mch <- haem %>% 
    select(id, date, mch) %>% 
    full_join(y = mch, by = c("id", "date")) %>% 
    mutate(mch = if_else(is.na(mch), value, mch)) %>% 
    select(id, date, mch) %>% 
    mutate(test = "mch") %>% 
    rename(value = mch) %>% 
    relocate(test, .before = value) %>% distinct()

mchc <- haem %>% 
    select(id, date, mchc) %>% 
    full_join(y = mchc, by = c("id", "date")) %>% 
    mutate(mchc = if_else(is.na(mchc), value, mchc)) %>% 
    select(id, date, mchc) %>% 
    mutate(test = "mchc") %>% 
    rename(value = mchc) %>% 
    relocate(test, .before = value) %>% distinct()

mcv <- haem %>% 
    select(id, date, mcv) %>% 
    full_join(y = mcv, by = c("id", "date")) %>% 
    mutate(mcv = if_else(is.na(mcv), value, mcv)) %>% 
    select(id, date, mcv) %>% 
    mutate(test = "mcv") %>% 
    rename(value = mcv) %>% 
    relocate(test, .before = value) %>% distinct()

wcc <- haem %>% 
    select(id, date, wbc) %>% 
    full_join(y = wcc, by = c("id", "date")) %>% 
    mutate(wbc = if_else(is.na(wbc), value, wbc)) %>% 
    select(id, date, wbc) %>% 
    mutate(test = "wcc") %>% 
    rename(value = wbc) %>% 
    relocate(test, .before = value) %>% distinct()

platelet <- haem %>% 
    select(id, date, plt) %>% 
    full_join(y = platelet, by = c("id", "date")) %>% 
    mutate(plt = if_else(is.na(plt), value, plt)) %>% 
    select(id, date, plt) %>% 
    mutate(test = "platelet") %>% 
    rename(value = plt) %>% 
    relocate(test, .before = value) %>% distinct()
    
dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "iron", na = "NULL", skip = 1,
                 col_names = c("id", "date", "ferritin", "iron",
                               "transferrin", "tsat"))
dt <- dt %>% 
    arrange(id, date)

ferritin <- dt %>% 
    select(id, date, ferritin) %>% 
    full_join(y = ferritin, by = c("id", "date")) %>% 
    mutate(ferritin = if_else(is.na(ferritin), value, ferritin)) %>% 
    filter(ferritin < 60000) %>% 
    select(id, date, ferritin) %>% 
    mutate(test = "ferritin") %>% 
    rename(value = ferritin) %>% 
    relocate(test, .before = value) %>% distinct()

iron <- dt %>% 
    select(id, date, iron) %>% 
    full_join(y = iron, by = c("id", "date")) %>% 
    mutate(iron = if_else(is.na(iron), value, iron)) %>% 
    select(id, date, iron) %>% 
    mutate(test = "iron") %>% 
    rename(value = iron) %>% 
    relocate(test, .before = value) %>% distinct()

transferrin <- dt %>% 
    select(id, date, transferrin) %>% 
    full_join(y = transferrin, by = c("id", "date")) %>% 
    mutate(transferrin = if_else(is.na(transferrin), value, transferrin)) %>% 
    select(id, date, transferrin) %>% 
    mutate(test = "transferrin") %>%
    rename(value = transferrin) %>% 
    relocate(test, .before = value) %>% distinct()

tsat <- dt %>% 
    select(id, date, tsat) %>% 
    full_join(y = tsat, by = c("id", "date")) %>% 
    mutate(tsat = if_else(is.na(tsat), value, tsat)) %>% 
    select(id, date, tsat) %>% 
    mutate(test = "tsat") %>% 
    rename(value = tsat) %>% 
    relocate(test, .before = value) %>% distinct()

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "calcium", na = "NULL", skip = 1,
                 col_names = c("id", "date", "calcium", "adjca",
                               "phosphate", "pth", "alkphos"))
dt <- dt %>% 
    arrange(id, date)

ca_adj <- dt %>% 
    select(id, date, adjca) %>% 
    full_join(y = ca_adj, by = c("id", "date")) %>% 
    mutate(adjca = if_else(is.na(adjca), value, adjca)) %>% 
    select(id, date, adjca) %>% 
    mutate(test = "adjca") %>% 
    rename(value = adjca) %>% 
    relocate(test, .before = value) %>% distinct() 

ca <- dt %>% 
    select(id, date, calcium) %>% 
    full_join(y = ca, by = c("id", "date")) %>% 
    mutate(value = if_else(is.na(value), calcium, value)) %>%
    filter(value < 10) %>%
    select(id, date, value) %>% 
    mutate(test = "calcium") %>% 
    relocate(test, .before = value) %>% distinct()

phosph <- dt %>% 
    select(id, date, phosphate) %>% 
    full_join(y = phosph, by = c("id", "date")) %>% 
    mutate(phosphate = if_else(is.na(phosphate), value, phosphate)) %>% 
    filter(phosphate < 10) %>% 
    select(id, date, phosphate) %>% 
    mutate(test = "phosphate") %>% 
    rename(value = phosphate) %>% 
    relocate(test, .before = value) %>% distinct()

pth <- dt %>% 
    select(id, date, pth) %>% 
    full_join(y = pth, by = c("id","date")) %>% 
    mutate(value = if_else(is.na(value), pth, value)) %>% 
    select(id, date, value) %>% 
    mutate(test = "pth") %>% 
    relocate(test, .before = value) %>% distinct() %>% drop_na(value)

alkphos <- dt %>% 
    select(id, date, alkphos) %>% 
    full_join(y = alkphos, by = c("id", "date")) %>% 
    mutate(value = if_else(is.na(value), alkphos, value)) %>% 
    select(id, date, value) %>% 
    mutate(test = "alkphos") %>% 
    relocate(test, .before = value) %>% distinct() %>% drop_na(value)

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "mag", na = "NULL", skip = 1,
                 col_names = c("id", "date", "magnesium"))
dt <- dt %>% 
    arrange(id, date)

mg <- dt %>% 
    full_join(y = mg, by = c("id", "date")) %>% 
    mutate(magnesium = if_else(is.na(magnesium), value, magnesium)) %>% 
    filter(magnesium < 10) %>% 
    select(id, date, magnesium) %>% 
    mutate(test = "magnesium") %>% 
    rename(value = magnesium) %>% 
    relocate(test, .before = value) %>% distinct()

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "electrolytes",na = "NULL", skip = 1,
                 col_names = c("id", "date", "sodium", "potassium",
                               "chloride", "bicarb", "creatinine", 
                               "urate", "urea"),
                 col_types = c("numeric", "date", "numeric", "numeric",
                               "numeric", "numeric", "numeric",
                               "numeric", "numeric"))
dt <- dt %>% 
    arrange(id, date)

sodium <- dt %>% 
    select(id, date, sodium) %>% 
    full_join(y = sodium, by = c("id", "date")) %>% 
    mutate(sodium = if_else(is.na(sodium), value, sodium)) %>% 
    filter(sodium != 5.8, sodium != 1238) %>% 
    select(id, date, sodium) %>% 
    mutate(test = "sodium") %>% 
    rename(value = sodium) %>% 
    relocate(test, .before = value) %>% distinct()

potassium <- dt %>% 
    select(id, date, potassium) %>% 
    full_join(y = potassium, by = c("id", "date")) %>% 
    mutate(potassium = if_else(is.na(potassium), value, potassium)) %>% 
    filter(potassium != 0, potassium < 10) %>% 
    select(id, date, potassium) %>% 
    mutate(test = "potassium") %>% 
    rename(value = potassium) %>% 
    relocate(test, .before = value) %>% distinct()

chloride <- dt %>% 
    select(id, date, chloride) %>% 
    full_join(y = chloride, by = c("id", "date")) %>% 
    filter(chloride > 70) %>% 
    select(id, date, chloride) %>% 
    mutate(test = "chloride") %>% 
    rename(value = chloride) %>% 
    relocate(test, .before = value) %>% distinct()

bicarb <- dt %>% 
    select(id, date, bicarb) %>% 
    full_join(y = bicarb, by = c("id", "date")) %>% 
    mutate(bicarb = if_else(is.na(bicarb), value, bicarb)) %>% 
    filter(bicarb < 50) %>% 
    select(id, date, bicarb) %>% 
    mutate(test = "bicarb") %>% 
    rename(value = bicarb) %>% 
    relocate(test, .before = value) %>% distinct() 

creatinine <- dt %>% 
    select(id, date, creatinine) %>% 
    full_join(y = creatinine, by = c("id", "date")) %>%
    mutate(value = if_else(is.na(value), creatinine, value),
           value = if_else(value < 10, value*1000, value)) %>% 
    select(id, date, value) %>% 
    mutate(test = "creatinine") %>% 
    relocate(test, .before = value) %>% distinct()

urate <- dt %>% 
    select(id, date, urate) %>% 
    full_join(y = urate, by = c("id", "date")) %>% 
    mutate(value = if_else(is.na(value), urate, value),
           value = if_else(value < 40, value*1000, value)) %>% 
    select(id, date, value) %>% 
    mutate(test = "urate") %>% 
    relocate(test, .before = value) %>% distinct() %>% drop_na(value)

urea <- dt %>% 
    select(id, date, urea) %>% 
    full_join(y = urea, by = c("id", "date")) %>%
    mutate(urea = if_else(is.na(urea), value, urea)) %>% 
    select(id, date, urea) %>% 
    mutate(test = "urea") %>% 
    rename(value = urea) %>% 
    relocate(test, .before = value) %>% distinct()

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "lfts", na = "NULL", skip = 1,
                 col_names = c("id", "date", "alkphos", "alt",
                               "albumin", "ggt", "bilirubin", "protein"),
                 col_types = c("numeric", "date", "skip", "numeric", 
                               "numeric", "numeric", "numeric", "numeric"))
dt <- dt %>% 
    arrange(id, date)

alt <- dt %>% 
    select(id, date, alt) %>% 
    full_join(y = alt, by = c("id", "date")) %>% 
    mutate(alt = if_else(is.na(alt), value, alt)) %>% 
    select(id, date, alt) %>% 
    mutate(test = "alt") %>% 
    rename(value = alt) %>% 
    relocate(test, .before = value) %>% distinct()

alb <- dt %>% 
    select(id, date, albumin) %>% 
    full_join(y = alb, by = c("id", "date")) %>% 
    mutate(albumin = if_else(is.na(albumin), value, albumin)) %>% 
    filter(between(albumin, 10, 100)) %>% 
    select(id, date, albumin) %>% 
    mutate(test = "albumin") %>% 
    rename(value = "albumin") %>% 
    relocate(test, .before = value) %>% distinct()

ggt <- dt %>% 
    select(id, date, ggt) %>% 
    full_join(y = ggt, by = c("id", "date")) %>% 
    mutate(value = if_else(is.na(value), ggt, value)) %>% 
    select(id, date, value) %>% 
    mutate(test = "ggt") %>% 
    relocate(test, .before = value) %>% distinct() 

bili <- dt %>% 
    select(id, date, bilirubin) %>% 
    full_join(y = bili, by = c("id", "date")) %>% 
    mutate(value = if_else(is.na(value), bilirubin, value)) %>% 
    select(id, date, value) %>% 
    mutate(test = "bilirubin") %>% 
    relocate(test, .before = value) %>% distinct() 

protein <- dt %>% 
    select(id, date, protein) %>% 
    full_join(y = protein, by = c("id", "date")) %>%
    mutate(value = if_else(is.na(value), protein, value)) %>% 
    select(id, date, value) %>% 
    mutate(test = "protein") %>% 
    relocate(test, .before = value) %>% distinct() %>% drop_na(value)

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "crp", skip = 1, na = "NULL",
                 col_names = c("id", "date", "crp"),
                 col_types = c("numeric", "date", "numeric"))
dt <- dt %>% 
    arrange(id, date)

crp <- dt %>% 
    full_join(y = crp, by = c("id", "date")) %>% 
    mutate(crp = if_else(is.na(crp), value, crp)) %>% 
    select(id, date, crp) %>% 
    mutate(test = "crp") %>% 
    rename(value = crp) %>% 
    relocate(test, .before = value) %>% distinct() 

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "globulin", skip = 1, na = "NULL",
                 col_names = c("id", "date", "globulin"))

globulin <- dt %>% 
    arrange(id, date) %>% 
    full_join(y = globulin, by = c("id", "date")) %>% 
    mutate(globulin = if_else(is.na(globulin), value, globulin)) %>% 
    select(id, date, globulin) %>% 
    mutate(test = "globulin") %>% 
    rename(value = globulin) %>% 
    relocate(test, .before = value) %>% distinct() 

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                  sheet = "egfr", skip = 1, na = "NULL",
                  col_names = c("id", "date", "egfrnum", "egfr", "creatinine"),
                  col_types = c("numeric", "date", "numeric", "text",
                                "numeric"))

dt <- dt %>% 
    arrange(id, date) %>% 
    mutate(egfr = parse_number(egfr)) %>% 
    select(-egfrnum)

egfr <- dt %>% 
    select(id, date, egfr) %>% 
    full_join(y = egfr, by = c("id", "date")) %>% 
    mutate(value = if_else(is.na(value), egfr, value)) %>% 
    filter(value <= 90) %>% 
    select(id, date, value) %>% 
    mutate(test = "egfr") %>% 
    relocate(test, .before = value) %>% 
    distinct() %>% 
    arrange(id, date)

dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "esr", skip = 1, na = "NULL",
                 col_names = c("id", "date", "esrnum", "esr"),
                 col_types = c("numeric", "date", "numeric", "text"))

dt <- dt %>% 
    arrange(id, date) %>% 
    mutate(esr = parse_number(esr)) %>% 
    select(-esrnum)

esr <- dt %>% 
    full_join(y = esr, by = c("id", "date")) %>% 
    mutate(esr = if_else(is.na(esr), value, esr)) %>% 
    select(id, date, esr) %>% 
    mutate(test = "esr") %>% 
    rename(value = esr) %>% 
    relocate(test, .before = value) %>% distinct()
```

```{r vital sign}
dt <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "Exams", na = "NULL", skip = 1,
                 col_types = c("numeric", "date", "numeric", "numeric",
                               "numeric", "numeric", "numeric", "numeric",
                               "numeric", "numeric"),
                 col_names = c("id", "date", "sitsys", "sitdias", "sithr",
                               "stasys", "stadias", "stahr", "weight",
                               "height"))

dt <- dt %>% 
    arrange(id, date)

sitsys <- dt %>% 
    select(id, date, sitsys) %>% 
    filter(between(sitsys, 60, 250)) %>% 
    mutate(test = "sitsys") %>% 
    rename(value = sitsys) %>% 
    relocate(test, .before = value) %>% distinct()

sitdias <- dt %>% 
    select(id, date, sitdias) %>% 
    filter(sitdias > 20) %>% 
    mutate(test = "sitdias") %>% 
    rename(value = sitdias) %>% 
    relocate(test, .before = value) %>% distinct()

sithr <- dt %>% 
    select(id, date, sithr) %>%
    mutate(test = "sithr") %>% 
    rename(value = sithr) %>% 
    relocate(test, .before = value) %>% distinct()

stasys <- dt %>% 
    select(id, date, stasys) %>% 
    filter(between(stasys, 50, 250)) %>% 
    mutate(test = "stasys") %>% 
    rename(value = stasys) %>% 
    relocate(test, .before = value) %>% distinct()

stadias <- dt %>% 
    select(id, date, stadias) %>% 
    filter(stadias > 20) %>% 
    mutate(test = "stadias") %>% 
    rename(value = stadias) %>% 
    relocate(test, .before = value) %>% distinct()

stahr <- dt %>% 
    select(id, date, stahr) %>% 
    filter(between(stahr, 30, 160)) %>%
    mutate(test = "stahr") %>% 
    rename(value = stahr) %>% 
    relocate(test, .before = value) %>% distinct()

weight <- dt %>% 
    select(id, date, weight) %>% 
    filter(between(weight, 20, 310)) %>% 
    mutate(test = "weight") %>% 
    rename(value = weight) %>% 
    relocate(test, .before = value) %>% distinct()
    
height <- dt %>% 
    select(id, date, height) %>% 
    filter(between(height, 50, 250)) %>% 
    mutate(test = "height") %>% 
    rename(value = height) %>% 
    relocate(test, .before = value) %>% distinct()
    
vital_signs <- na.omit(unique(rbindlist(
    list(sitsys, sitdias, sithr, stasys, stadias, stahr, weight, height)
)))

dt <- vital_signs %>% 
    pivot_wider(id_cols = c("id", "date"),
                names_from = test,
                values_from = value,
                values_fn = mean) %>% 
    mutate(height_m = height / 100,
           bmi = weight/(height_m^2),
           postural_sys = stasys - sitsys,
           postural_dias = stadias - sitdias,
           postural_hr = stahr - sithr)

bmi_posturals <- dt %>% 
    select(id, date, bmi:postural_hr) %>% 
    pivot_longer(cols = bmi:postural_hr,
                 names_to = "test",
                 values_to = "value")
```

```{r compilation of clinico-pathological data}
melted <- na.omit(unique(rbindlist(
    list(alb, albcreat, alkphos, alt, baso, bicarb, bili, ca, ca_adj, chloride, 
         creatinine, crp, egfr, eosin, esr, ferritin, ggt, globulin, 
         glucose, hb, hba1c, hct, iron, lymph, mch, mchc, mcv, mg, monocyte,
         neut, phosph, platelet, potassium, protcreat, protein, pth, rbc, rdw, 
         sodium, transferrin, tsat, urate, urea, wcc, vital_signs,
         bmi_posturals))))
saveRDS(melted, file = 'melted.RDS')
```
Extracting from lab_results (raw text), there were 10577 distinct IDs. While, from pathology, there were 10,564 distinct IDs. From the eGFR perspective (new extraction), there were 9,700 distinct IDs.
We had extracted 15,481 distinct IDs while Melted had 10198 distinct IDs (based on test == "egfr"). 
```{r demographic data and modality}
patient <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "IDs", na = "NULL",
                 col_names = c("id", "birth_year", "sex"),
                 col_types = c("numeric", "numeric", "text"),
                 skip = 1)
patient <- patient %>% distinct(id, .keep_all = TRUE)

dt <- read_excel("~/PhD Project 1/ML for ESKD/need_gender_clarification.xlsx",
                 sheet = "Sheet1", na = "NULL",
                 col_names = c("id", "sex"),
                 col_types = c("numeric", "text"),
                 skip = 1)
patient <- patient %>% 
    left_join(y = dt, by = "id") %>% 
    mutate(sex.x = case_when(
        sex.x == "Unknown" ~ sex.y,
        sex.x == "Indeterminate" ~ sex.y,
        TRUE ~ sex.x
    )) %>% 
    select(id, birth_year, sex.x) %>% 
    rename(sex = sex.x) %>% 
    arrange(id)

dt <- read_excel("~/PhD Project 1/ML for ESKD/Diagnosis.xlsx",
                 sheet = "GN Diagnosis", skip = 1, na = "NULL",
                 col_names = c("id", "diagnosis", "diag_date"),
                 col_types = c("numeric", "text", "date"))

lupus <- tibble(unified_diag = "lupus", 
              diagnosis = c('Systemic lupus erythematosus glomerulonephritis syndrome', 
                            'Systemic lupus erythematosus glomerulonephritis syndrome, World Health Organization (WHO) class III', 
                            'Systemic lupus erythematosus glomerulonephritis syndrome, World Health Organization (WHO) class IV', 
                            'Systemic lupus erythematosus glomerulonephritis syndrome, World Health Organization (WHO) class V', 
                            'Systemic lupus erythematosus glomerulonephritis syndrome, World Health Organization (WHO) class VI'))

unspecific <- tibble(unified_diag = "unspecific",
              diagnosis = c('Acute glomerulonephritis', 
                            'Chronic diffuse glomerulonephritis',
                            'Chronic focal glomerulonephritis', 
                            'Chronic glomerulonephritis',
                            'Crescentic glomerulonephritis',
                            'Focal AND segmental proliferative glomerulonephritis',
                            'Glomerulonephritis', 
                            'Immune-complex glomerulonephritis',
                            'Necrotizing glomerulonephritis',
                            'Proliferative glomerulonephritis', 
                            'Renal vasculitis',
                            'Vasculitis secondary to drug',
                            'Acute post-streptococcal glomerulonephritis',
                            'Post-infectious glomerulonephritis'))

mp_mcgn <- tibble(unified_diag = "mp_mcgn",
              diagnosis = c('Acute nephritic syndrome, diffuse mesangial proliferative glomerulonephritis', 
                            'C3 Glomerulonephritis',
                            'Mesangial proliferative glomerulonephritis',
                            'Mesangial proliferative glomerulonephritis',
                            'Mesangiocapillary glomerulonephritis',
                            'Mesangiocapillary glomerulonephritis, type I',
                            'Mesangiocapillary glomerulonephritis, type II'))

anca <- tibble(unified_diag = "anca",
              diagnosis = c('Eosinophilc Granulomatosis with Polyangiitis',
                            'Granulomatous Polyangiitis', 
                            'Microscopic polyangiitis',
                            'Microscopic Polyangiitis',
                            'Primary pauci-immune necrotizing and crescentic glomerulonephritis',
                            "Wegener's granulomatosis"))
fg_itg <- tibble(unified_diag = "fg_itg",
              diagnosis = c('Fibrillary glomerulonephritis', 
                            'Immunotactoid Glomerulopathy'))

fsgs <- tibble(unified_diag = "fsgs",
              diagnosis = c('Focal Segmental Glomerulosclerosis'))
igan <- tibble(unified_diag = "igan",
              diagnosis = c('IgA nephropathy', 'Primary IgA nephropathy'))
mn <- tibble(unified_diag = "mn",
              diagnosis = c('Membranous glomerulonephritis',
                            'Membranous glomerulonephritis - stage III',
                            'Nephrotic syndrome, diffuse membranous glomerulonephritis')) 
mcd <- tibble(unified_diag = "mcd",
               diagnosis = c('Minimal change disease',
                             'Steroid-dependent minimal change glomerulonephritis'))

gn_reference <- rbindlist(list(lupus, unspecific, mp_mcgn, anca, fg_itg, 
                               fsgs, igan, mn, mcd))

gn <- dt %>% 
    select(-diag_date) %>% 
    left_join(gn_reference, by = "diagnosis") %>% 
    select(-diagnosis) %>% 
    rename(diagnosis = unified_diag) %>% 
    arrange(id) %>% drop_na()

gn <- gn %>% 
    mutate(diagnosis = replace(diagnosis, 
                               id == 4339 & diagnosis == "mcd", NA),
           diagnosis = replace(diagnosis, 
                               id == 4551 & diagnosis == "mn", NA),
           diagnosis = replace(diagnosis, 
                               id == 10143 & diagnosis == "igan", NA),
           diagnosis = replace(diagnosis, 
                               id == 15613 & diagnosis == "mcd", NA),
           diagnosis = replace(diagnosis, 
                               id == 18367 & diagnosis == "unspecific", 
                               NA)) %>%
    drop_na() %>% 
    distinct(id, .keep_all = TRUE) 

death <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "Mortality", na = "NULL",
                 col_names = c("id", "death_date"),
                 col_types = c("numeric", "date"),
                 skip = 1)
death <- death %>% drop_na()

modality <- read_excel("~/PhD Project 1/ML for ESKD/new_pathology.xlsx",
                 sheet = "Modality", na = "NULL",
                 col_names = c("id", "start", "stop", "modality"),
                 col_types = c("numeric", "date", "date", "text"),
                 skip = 1)
modality <- modality %>% 
    arrange(id, start)

eskd_reference <- c("Haemodiafiltration", "Haemofiltration", "HD",
                    "Nocturnal HD", "PD", "Transplant")

eskd_event <- modality %>% 
    filter(modality %in% eskd_reference) %>% 
    group_by(id) %>% 
    arrange(start) %>% 
    slice_head() %>% 
    ungroup() %>% 
    select(id, start, modality)
```
There were 15 IDs without baseline demographics (sex and birthyear) and 500 IDs with pre-existing ESKD (Hx of ESKD before initial eGFR). Hence, a total of 9683 IDs for outcome_dt
```{r outcome and feature dataset}
initial_egfr <- melted %>%
    filter(test == "egfr") %>% 
    group_by(id) %>% 
    arrange(date) %>% 
    slice_head() %>% ungroup() %>% 
    mutate(test = "egfr_init")

last_egfr <- melted %>%
    filter(test == "egfr") %>% 
    group_by(id) %>% 
    arrange(date) %>% 
    slice_tail() %>% ungroup() %>% 
    mutate(test = "egfr_last")

egfr_15 <- melted %>% 
    filter(test == "egfr") %>% 
    group_by(id) %>% 
    arrange(date) %>% 
    filter(value <= 15) %>% 
    slice_head() %>% 
    ungroup() %>% 
    mutate(test = "egfr_eskd")

egfr_melt <- na.omit(unique(rbindlist(list(initial_egfr, last_egfr,
                                      egfr_15))))

dt <- egfr_melt %>% 
    pivot_wider(id_cols = id,
                names_from = test,
                values_from = c("date", "value")) %>% 
    left_join(y = patient, by = ("id")) %>% 
    drop_na(birth_year) %>% 
    mutate(age_init = year(date_egfr_init) - birth_year) %>% 
    left_join(y = death, by = c("id")) %>% 
    left_join(y = eskd_event, by = c("id")) 
    
dt <- dt %>% 
    mutate(preexisting_eskd = 
               if_else(start <= date_egfr_init, 
                       "yes", "no", missing = "no")) %>% 
    filter(preexisting_eskd == "no") %>% 
    select(-preexisting_eskd)

dt <- dt %>% 
    mutate(modality_corrected_eskd_date = if_else(
        date_egfr_eskd > start & !is.na(date_egfr_eskd) & !is.na(start), 
        start, date_egfr_eskd
    )) %>% 
    mutate(recovery = if_else(
        is.na(modality) & value_egfr_last >= 17 & !is.na(date_egfr_eskd),
        "yes",
        "no"
    )) 

outcome_dt <- dt %>% 
    mutate(modality_corrected_eskd_date = replace(modality_corrected_eskd_date,
                                                  recovery == "yes",
                                                  NA)) %>% 
    select(-date_egfr_last, -date_egfr_eskd, -(value_egfr_last:birth_year), 
           -start, -modality, -recovery) %>% 
    rename(eskd_date = modality_corrected_eskd_date) %>% 
    relocate(eskd_date, .before = death_date)

dt <- outcome_dt %>%
    select(id, date_egfr_init) %>%
    right_join(y = melted, by = c("id")) %>% 
    as.data.table()

dt[, ":="(resample_date = floor_date(date, unit = "month"),
          resample_egfr_init = floor_date(date_egfr_init, unit = "month"))] %>% 
    .[, by = c("id", "test", "resample_date"), monthly_value := mean(value)]

feature_dt <- dt %>% 
    select(id, resample_date, resample_egfr_init, test, monthly_value) %>% 
    distinct() %>% 
    mutate(relyear = as.duration(resample_date-resample_egfr_init)/dyears(1)) %>% 
    filter(relyear >= 0) %>% 
    select(id, test, monthly_value, relyear) %>% 
    arrange(id, relyear) %>% 
    rename(value = monthly_value)
# fwrite(outcome_dt, "outcome_data.gz")
# fwrite(feature_dt, "feature_data_resample.gz")
```
There were 577 IDs who developed ESKD within 0.2 years with initial eGFR (median 10, IQR 6.54 to 13.5) and 2 IDs who died less than 0.01 years (~3 days). Hence, a total of 9104 IDs for further analysis
```{r meta dataset}
dt <- outcome_dt %>% 
    mutate(follow_up = as.duration(extraction_date - date_egfr_init)/dyears(1),
           eskd_fu = as.duration(eskd_date - date_egfr_init)/dyears(1),
           death_fu = as.duration(death_date - date_egfr_init)/dyears(1)) %>% 
    filter(death_fu > 0 | is.na(death_fu), 
           eskd_fu > 0.2 | is.na(eskd_fu)) %>% 
    mutate(time_eskd = case_when(
        is.na(eskd_fu) & is.na(death_fu) ~ follow_up,
        is.na(eskd_fu) & !is.na(death_fu) ~ death_fu,
        TRUE ~ eskd_fu
    ),
    status_eskd = if_else(
        time_eskd == eskd_fu, "eskd", "no", missing = "no"
    )) %>% 
    mutate(time_compete = pmin(follow_up, eskd_fu, death_fu, na.rm = TRUE),
           status_compete = case_when(
               time_compete == eskd_fu ~ "eskd",
               time_compete == death_fu ~ "death",
               time_compete == follow_up ~ "censored"
           ))

meta_dt <- dt %>% 
    left_join(y = gn, by = c("id")) %>% 
    rename(gn = diagnosis) %>% 
    relocate(gn, .after = age_init) %>% 
    replace_na(list(gn = "no")) %>% 
    mutate(gn_cat = if_else(gn == "no", "no", "yes")) %>% 
    relocate(gn_cat, .after = gn) %>% 
    left_join(select(feature_dt, id:relyear), by = c("id")) %>% 
    relocate(test:relyear, .after = id) %>% 
    filter(relyear < time_compete)
saveRDS(meta_dt, "meta.RDS")
```

```{r counting clinico-pathology frequency}
test_list <- meta_dt %>%
    select(id, test:relyear) %>% 
    count(test) %>% 
    pull(test)

id_dt <- meta_dt %>% 
    select(id) %>% 
    distinct() %>% as.data.table()

dt <- meta_dt %>%
    select(id:relyear) %>% 
    as.data.table()

for (test_name in test_list){
    m <- dt[test == test_name]
    n <- m[, .N, by = id]
    setnames(n, old = 'N', new = test_name)
    setkey(n, id)
    id_dt <- n[id_dt]
}
```
From 9104 IDs, only 5315 IDs had at least 3 eGFR. Of this 5315, 845 (15.89%) had ESKD and 916 (17.23%) died. 

When we enforced additional restriction on 5315 IDs:
Dense3 had 4263 IDs with 607 (14.23%) ESKD and 716 (16.79%) death.
Tangri3 had a total 2685 IDs with 294 (10.94%) ESKD and 283 (10.54%) death.
```{r creation of dataset}
egfr3_id <- id_dt[egfr >=3, id]

dense3_id <- id_dt %>% 
    filter(egfr >= 3) %>% 
    select(id, albumin, alkphos, bicarb, calcium, chloride, glucose, 
           haemoglobin, phosphate, platelet, potassium, sodium, urea, wcc) %>% 
    filter(if_all(.fns = ~ . >=3)) %>% 
    pull(id)

tangri3_id <- id_dt %>% 
    filter(egfr >= 3) %>% 
    select(id, albcreat_ratio, albumin, phosphate, bicarb, adjca) %>% 
    filter(if_all(.fns = ~. >=3)) %>% 
    pull(id)

dt <- meta_dt %>% 
    filter(id %in% dense3_id) %>% 
    select(-date_egfr_init)
saveRDS(dt, "dense3.RDS")

# dt <- meta_dt %>% 
#     filter(id %in% tangri3_id) %>% 
#     select(-date_egfr_init)
# saveRDS(dt, "tangri3.RDS")
```

```{r function to assess outcome based on filtered ids}
events_freq <- function(id){
    meta_dt %>% 
        filter(id %in% {{id}}) %>% 
        select(id, status_compete) %>% 
        distinct() %>% 
        count(status_compete)
}

events_freq(egfr3_id)
```

```{r heatmap of data density, eval=FALSE}
dt <- id_dt[id %chin% dense3_id,]
dt[, id:=NULL]
dt[is.na(dt)] <- 0    

heatmap_breaks<-c(0, 1, 2, 5, 10, 20, 50, 100, 200, 500, 1000, 2000, 5000, 10000)
heatmaply(x = t(as.matrix(dt)), ylab = 'Features', xlab = 'Patients', 
          main = 'Data Density of Clinico-Pathological Data of Patients in Dense3 IDs',
          dist_method = 'manhattan', hclust_method = 'ward.D2',
          Rowv = TRUE, Colv = TRUE, dendrogram = 'both', 
          showticklabels = c(FALSE, TRUE), 
          scale_fill_gradient_fun = scale_fill_viridis(trans='sqrt', 
                                                       breaks=heatmap_breaks,
                                                       labels=heatmap_breaks, 
                                                       option = 'B', 
                                                       direction = -1,
                                                       na.value='white'))
```

